\chapter{Lorsque la poésie justifie l'étude de la phonétique}

Pour écouter ce poème rendez-vous sur cet
\href{http://doyouspeakenglish.fr/un-poeme-qui-justifie-letude-de-la-phonetique/}{article}
de mon blog.

\begin{itemize}
\item \exEN{Ration never rhymes with nation,}
\item \textcolor{teal}{ˈræʃən ˈnɛvə raɪmz wɪð ˈneɪʃən,}
\item \exEN{Say prefer, but preferable,}
\item \textcolor{teal}{seɪ priˈfɜː, bʌt ˈprɛfərəbl,}
\item \exEN{Comfortable and vegetable.}
\item \textcolor{teal}{ˈkʌmfətəbl ænd ˈvɛʤɪtəbl.}
\item \exEN{B must not be heard in doubt,}
\item \textcolor{teal}{biː mʌst nɒt biː hɜːd ɪn daʊt,}
\item \exEN{Debt and dumb both leave it out, }
\item \textcolor{teal}{dɛt ænd dʌm bəʊθ liːv ɪt aʊt,}
\item \exEN{In the words psychology,}
\item \textcolor{teal}{ɪn ðə wɜːdz saɪˈkɒləʤi,}
\item \exEN{Psychic, and psychiatry,}
\item \textcolor{teal}{ˈsaɪkɪk, ænd saɪˈkaɪətri,}
\item \exEN{You must never sound the p.}
\item \textcolor{teal}{juː mʌst ˈnɛvə saʊnd ðə piː.}
\item \exEN{Psychiatrist you call the man}
\item \textcolor{teal}{saɪˈkaɪətrɪst juː kɔːl ðə mæn}
\item \exEN{Who cures the complex, if he can.}
\item \textcolor{teal}{huː kjʊəz ðə ˈkɒmplɛks, ɪf hiː kæn.}
\item \exEN{In architect, chi is k.}
\item \textcolor{teal}{ɪn ˈɑːkɪtɛkt, ʧiː ɪz keɪ.}
\item \exEN{In arch it is the other way.}
\item \textcolor{teal}{ɪn ɑːʧ ɪt ɪz ði ˈʌðə weɪ.}
\item \exEN{Please remember to say iron}
\item \textcolor{teal}{pliːz rɪˈmɛmbə tə seɪ ˈaɪən}
\item \exEN{So that it'll rhyme with lion.}
\item \textcolor{teal}{səʊ ðət ˈɪtl raɪm wɪð ˈlaɪən.}
\item \exEN{Advertisers advertse,}
\item \textcolor{teal}{ˈædvətaɪzəz advertse,}
\item \exEN{Advertisements will put you wise.}
\item \textcolor{teal}{ədˈvɜːtɪsmənts wɪl pʊt jʊ waɪz.}
\item \exEN{Time when work is done is leisure,}
\item \textcolor{teal}{taɪm wɛn wɜːk s dʌn z ˈlɛʒə,}
\item \exEN{Fill it up with useful pleasure.}
\item \textcolor{teal}{fɪl ɪt ʌp wɪð ˈjuːsfʊl ˈplɛʒə.}
\item \exEN{Accidental, accident,}  
\item \textcolor{teal}{ˌæksɪˈdɛntl, ˈæksɪdənt,}
\item \exEN{Sound the g in ignorant.}
\item \textcolor{teal}{saʊnd ðə ʤiː ɪn ˈɪgnərənt.}
\item \exEN{Relative, but relation,}
\item \textcolor{teal}{ˈrɛlətɪv, bət rɪˈleɪʃən,}
\item \exEN{Then say creature, creation.}
\item \textcolor{teal}{ðɛn seɪ ˈkriːʧə, kri(ː)ˈeɪʃən.}
\item \exEN{Say the a in gas quite short,}
\item \textcolor{teal}{seɪ ði ə ɪn gæs kwaɪt ʃɔːt,}
\item \exEN{Bought remember rhymes with thwart,}
\item \textcolor{teal}{bɔːt rɪˈmɛmbə raɪmz wɪð θwɔːt,}
\item \exEN{Drought must always rhyme with bout,}
\item \textcolor{teal}{draʊt məst ˈɔːlweɪz raɪm wɪð baʊt,}
\item \exEN{In daughter leave the "gh" out.}
\item \textcolor{teal}{ɪn ˈdɔːtə liːv ðiː "gh" aʊt.}
\item \exEN{Wear a boot upon your foot.}
\item \textcolor{teal}{weər ə buːt əpən jə fʊt.}
\item \exEN{Root can never rhyme with soot.}
\item \textcolor{teal}{ruːt kən ˈnɛvə raɪm wɪð sʊt.}
\item \exEN{In muscle, sc is s,}
\item \textcolor{teal}{ɪn ˈmʌsl, sc z ɛs,}
\item \exEN{In muscular, it's sk, yes!}
\item \textcolor{teal}{ɪn ˈmʌskjʊlə, ɪts sk, jɛs!}
\item \exEN{Choir must always rhyme with wire,}
\item \textcolor{teal}{ˈkwaɪə məst ˈɔːlweɪz raɪm wɪð ˈwaɪə,}
\item \exEN{That again will rhyme with liar.}
\item \textcolor{teal}{ðæt əˈgɛn wɪl raɪm wɪð ˈlaɪə.}
\item \exEN{Then remember it's address.}
\item \textcolor{teal}{ðɛn rɪˈmɛmbər ɪts əˈdrɛs.}
\item \exEN{With an accent like posses.}
\item \textcolor{teal}{wɪð ən ˈæksənt laɪk ˈpɒsiz.}
\item \exEN{G in sign must be silent be,}
\item \textcolor{teal}{ʤiː ɪn saɪn məst bi ˈsaɪlənt biː,}
\item \exEN{In signature, pronounce the g.}
\item \textcolor{teal}{ɪn ˈsɪgnɪʧə, prəˈnaʊns ðə ʤiː.}
\item \exEN{Please remember, say towards}
\item \textcolor{teal}{pliːz rɪˈmɛmbə, seɪ təˈwɔːdz}
\item \exEN{Just as if it ryhmed with boards.}
\item \textcolor{teal}{dʒəst əz ɪf ɪt raɪmd wɪð bɔːdz}
\item \exEN{Weight's like wait, but not like height.}
\item \textcolor{teal}{weɪts laɪk weɪt, bət nɒt laɪk haɪt.}
\item \exEN{Which should always rhyme with might.}
\item \textcolor{teal}{wɪʧ ʃəd ˈɔːlweɪz raɪm wɪð maɪt.}
\item \exEN{Sew is just the same as so,}
\item \textcolor{teal}{səʊ z ʤəst ðə seɪm əz səʊ,}
\item \exEN{Tie a ribbon in a bow.}
\item \textcolor{teal}{taɪ ə ˈrɪbən ɪn ə baʊ.}
\item \exEN{When You meet the queen you bow,}
\item \textcolor{teal}{wɛn jʊ miːt ðə kwiːn jʊ baʊ,}
\item \exEN{Which again must rhyme with how.}
\item \textcolor{teal}{wɪʧ əˈgɛn məst raɪm wɪð haʊ.}
\item \exEN{In perfect English make a start.}
\item \textcolor{teal}{ɪn ˈpɜːfɪkt ˈɪŋglɪʃ meɪk ə stɑːt.}
\item \exEN{Learn this little rhyme by heart.}
\item \textcolor{teal}{lɜːn ðɪs ˈlɪtl raɪm baɪ hɑːt.}                    
\end{itemize}
